### 2015-11-01 per-coordinate.tex

\section{Per-Coordinate Learning}

An interesting class of algorithms proposed in~\cite{McMahan-Streeter-2010} and
\cite{Duchi-Hazan-Singer-2011} are based on the so-called per-coordinate
learning rates.  As shown in \cite{Streeter-McMahan-2010}, our algorithms, or
in fact any algorithm for OLO, can be used with per-coordinate
learning rates as well.

Abstractly, we assume that the decision set is a Cartesian product $K=K_1
\times K_2 \times \dots \times K_d$ of a finite number of convex sets.  On each
factor $K_i$, $i=1,2,\dots,d$, we can run any OLO algorithm separately and we
denote by $\Regret_T^{(i)}(u_i)$ its regret with respect to $u_i \in K_i$. The
overall regret with respect to any $u=(u_1, u_2, \dots, u_d) \in K$ can be
written as
$$
\Regret_T(u) = \sum_{i=1}^d \Regret_T^{(i)}(u_i) \; .
$$
If the algorithm for each factor is scale-free, the overall algorithm is
clearly scale-free as well.  Using \textsc{AdaFTRL} or \textsc{SOLO FTRL} for
each factor $K_i$, we generalize and improve existing regret bounds
\cite{McMahan-Streeter-2010,Duchi-Hazan-Singer-2011} for algorithms with per-coordinate
learning rates.