### 2015-11-23 solo-ftrl-proofs.tex

\section{Proofs for SOLO FTRL}
\label{section:solo-ftrl-proofs}

\begin{proof}[Proof of Lemma~\ref{lemma:sum-of-square-roots-inverses}]
We use the inequality $x/\sqrt{x+y} \le 2(\sqrt{x+y} - \sqrt{y})$ which holds
for non-negative $x,y$ that are not both zero. Substituting $x = a_t$ and $y=\sum_{s=1}^{t-1} a_s$,
we get that for any $t \ge 1$,
$$
\frac{a_t}{\sqrt{\sum_{s=1}^t a_s}} \le 2 \sqrt{\sum_{s=1}^t a_s} \ - \ 2 \sqrt{\sum_{s=1}^{t-1} a_s} \; .
$$
Summing the above inequality over all $t=1,2,\dots,T$, the right side telescopes to
$2 \sqrt{\sum_{t=1}^T a_t}$.
\end{proof}

\begin{proof}[Proof of Corollary~\ref{corollary:regret-solo-ftrl-bounded-set}]
Let $S = \sup_{v \in K} f(v)$. Theorem~\ref{theorem:regret-solo-ftrl} applied
to the regularizer $R(w) = \frac{c}{\sqrt{S}} f(w)$, together with
Proposition~\ref{proposition:diameter-vs-range} and a crude bound
$\max_{t=1,2,\dots,T} \|\ell_t\|_* \le \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2}$,
give
$$
\Regret_T \le \left(c + \frac{2.75}{c}  + 3.5\sqrt{8} \right) \sqrt{S \sum_{t=1}^{T} \norm{\ell_t}_*^2} \; .
$$
We choose $c$ by minimizing $g(c) = c + \frac{2.75}{c} + 3.5\sqrt{8}$. Clearly,
$g(c)$ has minimum at $c = \sqrt{2.75}$ and has minimal value $g(\sqrt{2.75}) =
2\sqrt{2.75} + 3.5\sqrt{8} \le 13.3$.
\end{proof}