### 2015-12-08 abstract.tex
Reworked introduction

We design algorithms for online linear optimization that have optimal regret
and at the same time do not need to know any upper or lower bounds on the norm
of the loss vectors.  We achieve adaptiveness to norms of loss vectors by scale
invariance, i.e., our algorithms make exactly the same decisions if the
sequence of loss vectors is multiplied by any positive constant.  Our
algorithms work for any decision set, bounded or unbounded.  For unbounded
decisions sets, these are the first adaptive algorithms for online linear
optimization.

We also study a popular scale-free variant of online mirror descent algorithm,
and we show that in two natural settings it has linear or worse regret.