### 2015-12-21 abstract.tex
Polished grammar and typography.

We design algorithms for online linear optimization that have optimal regret
and at the same time do not need to know any upper or lower bounds on the norm
of the loss vectors.  We achieve adaptiveness to the norms of the loss vectors
by scale invariance, i.e., our algorithms make exactly the same decisions if
the sequence of loss vectors is multiplied by any positive constant.  One of
our algorithms works for any decision set, bounded or unbounded.  For unbounded
decisions sets, it is the first adaptive algorithm for online linear
optimization with a non-vacuous bound.

We also study a popular scale-free variant of online mirror descent algorithm,
and we show that in two natural settings it has linear or worse regret.