### 2016-09-15 revision/abstract.tex
changes for reviewer 1

We design algorithms for online linear optimization that have optimal regret
and at the same time do not need to know any upper or lower bounds on the norm
of the loss vectors. We design them as instantiation of the Follow the Regularized Leader (FTRL) and Mirror Descent (MD) meta-algorithms.
We achieve adaptiveness to the norms of the loss vectors
by scale invariance, i.e., our algorithms make exactly the same decisions if
the sequence of loss vectors is multiplied by any positive constant.  The algorithm based on FTRL works for any decision set, bounded or unbounded.  For unbounded decisions sets, this is the first adaptive algorithm for online linear
optimization with a non-vacuous regret bound.
We also show lower bounds on scale-free algorithms based on MD on unbounded domains.