### 2016-09-15 revision/solo-ftrl-proofs.tex

\section{Proofs for \textsc{SOLO FTRL}}
\label{section:solo-ftrl-proofs}

\begin{proof}[Proof of Lemma~\ref{lemma:sum-of-square-roots-inverses}]
We use the inequality $x/\sqrt{x+y} \le 2(\sqrt{x+y} - \sqrt{y})$ which holds
for non-negative $x,y$ that are not both zero. Substituting $x = a_t$ and $y=\sum_{i=1}^{t-1} a_i$,
we get that for any $t \ge 1$,
$$
\frac{a_t}{\sqrt{\sum_{i=1}^t a_i}} \le 2 \sqrt{\sum_{i=1}^t a_i} \ - \ 2 \sqrt{\sum_{i=1}^{t-1} a_i} \; .
$$
Summing the above inequality over all $t=1,2,\dots,T$, the right side telescopes to
$2 \sqrt{\sum_{t=1}^T a_t}$.
\end{proof}