### 2016-09-27 revision/abstract.tex
Improvements

We design and analyze algorithms for online linear optimization that have
optimal regret and at the same time do not need to know any upper or lower
bounds on the norm of the loss vectors. Our algorithms are instances of the
Follow the Regularized Leader (FTRL) and Mirror Descent (MD) meta-algorithms. We
achieve adaptiveness to the norms of the loss vectors by scale invariance, i.e.,
our algorithms make exactly the same decisions if the sequence of loss vectors
is multiplied by any positive constant.  The algorithm based on FTRL works for
any decision set, bounded or unbounded.  For unbounded decisions sets, this is
the first adaptive algorithm for online linear optimization with a non-vacuous
regret bound. In contrast, we show lower bounds on scale-free algorithms based
on MD on unbounded domains.