### 2016-09-27 revision/conclusions.tex
Fix conclusions section

\section{Conclusions}
\label{section:conclusions}

We have investigated scale-free algorithms for online linear optimization and
we have shown that the scale-free property leads to algorithms which have optimal
regret and do not need to know or assume \textbf{anything} about the sequence
of loss vectors. In particular, the algorithms do not assume any upper or lower
bounds on the norms of the loss vectors or the number of rounds.

We have designed a scale-free algorithm based on \textsc{Follow The Regularizer
Leader}. Its regret is $O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} +
\min{\sqrt{T}, D} \max_{t=1,2,\dots,T} \|\ell_t\|_* \right)$ with respect to any
competitor $u \in K$, where $f$ is any non-negative $1$-strongly convex function
defined on $K$ and $D$ is the diameter of $K$. The result makes sense even when
the decision set $K$ is unbounded.

A similar, but weaker result holds for a scale-free algorithm based on
\textsc{Mirror Descent}. However, we have also shown that this algorithm is
strictly weaker than algorithms based on \textsc{Follow The Regularizer Leader}.
Namely, we gave examples of regularizers for which the scale-free version of
\textsc{Mirror Descent} has $\Omega(T)$ regret or worse.

We have proved an $\frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T \|\ell\|_*^2}$ lower
bound on the regret of any algorithm for any decision set with diameter $D$.

Notice that with the regularizer $f(u) = \frac{1}{2}\|u\|_2^2$ the regret of
\textsc{SOLO FTRL} depends quadratically on the norm of the competitor
$\|u\|_2$. There exist non-scale-free algorithms \cite{McMahan-Streeter-2012,
McMahan-Abernethy-2013, Orabona-2013, McMahan-Orabona-2014, Orabona-2014} that
have only a $O(\|u\|_2 \sqrt{\log \|u\|_2})$ or $O(\|u\|_2 \log \|u\|_2)$
dependency.  These algorithms assume an a priori bound on the norm of the loss
vectors. It remains an open problem to design algorithms that have sub-quadratic
dependency on $\norm{u}_2$ and are adaptive to the scale of the loss vectors.