### 2016-12-14 first-submission/ada-ftrl-proofs.tex
Rename directory

\section{Proofs for \textsc{AdaFTRL}}
\label{section:ada-ftrl-proofs}

\begin{proof}[Proof of Corollary~\ref{corollary:ada-ftrl-regret-bound}]
Let $S = \sup_{v \in K} f(v)$. Theorem~\ref{theorem:ada-ftrl-regret-bound}
applied to the regularizer $R(w) = \frac{c}{S} f(w)$ and
Proposition~\ref{proposition:diameter-vs-range} gives
$$
\Regret_T \le \sqrt{3}(1 + c) \max\left\{\sqrt{8}, \frac{1}{\sqrt{2c}} \right\} \sqrt{S \sum_{t=1}^T \|\ell_t\|_*^2} \; .
$$
It remains to find the minimum of $g(c) = \sqrt{3}(1 + c) \max\{\sqrt{8},
1/\sqrt{2c}\}$.  The function $g$ is strictly convex on $(0, \infty)$ and has
minimum at $c=1/16$ and $g(\frac{1}{16}) = \sqrt{3}(1+\frac{1}{16})\sqrt{8} \le
5.3$.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lemma:recurrence-solution}]
We prove the lemma by induction on $T$.  The base case $T=0$ is trivial.
For $T \ge 1$, let $f(x) = x + \min\{a_T, a_T^2/x\}$. We have
\begin{align*}
\Delta_T
& \le \Delta_{T-1} + \min \left\{a_T, \ \frac{a_T^2}{\Delta_{T-1}} \right\} \\
& = f \left( \Delta_{T-1} \right) \\
& \le f \left( \sqrt{3 \sum_{t=1}^{T-1} a_t^2} \right) \\
& = \sqrt{3 \sum_{t=1}^{T-1} a_t^2} \ + \ \min \left\{ a_T, \ \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}  \; ,
\end{align*}
where the second inequality follows from the inductive hypothesis $\Delta_{T-1} \le
\sqrt{3 \sum_{t=1}^{T-1} a_t^2}$ and that $f(x)$ is non-positive on $[-\infty,
0)$ and non-negative increasing on $[0,\infty)$. It remains to prove that
$$
\sqrt{3 \sum_{t=1}^{T-1} a_t^2} \ + \ \min \left\{ a_T, \ \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}
\le  \sqrt{3 \sum_{t=1}^T a_t^2} \; .
$$
If $a_T = 0$, the inequality clearly holds. Otherwise, $a_T$ is positive and we
divide through by it and make substitution $z=\frac{\sqrt{\sum_{t=1}^{T-1}
a_t^2}}{a_T}$, which leads to
$$
z\sqrt{3} + \min\left\{1, \frac{1}{z\sqrt{3}}\right\} \le \sqrt{3 + 3z^2}
$$
which can be easily checked by considering separately the cases $z \in
[0,\frac{1}{\sqrt{3}})$ and $z \in [\frac{1}{\sqrt{3}}, \infty)$.
\end{proof}